Air cargo load and route planning in pickup and delivery operations

### 2023-06-08 CAIE/cie.tex
cie file updated and experimentations almost ended

\usepackage{textcomp}
\usepackage{graphicx} 
\setlength{\fboxsep}{0.005pt}
\newcommand{\tmpframe}[1]{\fbox{#1}}

@@

As can be seen, so far Lurkin and Schyns \cite{LurkinSchyns2015} is the only work that simultaneously addresses an air cargo (WBP) and a flight itinerary (PDP) sub-problem. Although it is innovative, strong simplifications were imposed by the authors: in relation to loading, APP was ignored; with regard to routing, it is assumed a predefined tour plan restricted to two legs. It is important to note that these authors consider an aircraft with two doors, and the minimization of loading and unloading costs at the intermediate node was modeled through a container sequencing problem. Referring directly to this work, Brandt and Nickel \cite[p. 409]{BrandtStefan2019} comment: {\it However, not even these sub-problems are acceptably solved for real-world problem sizes or the models omit some practically relevant constraints}.

@@

However, as reported by Fok and Chun, {\it load planning (...) is usually done in roughly 2 hours before departure, when all the details of the cargo are present.} \cite{fok2004optimizing}. We may infer that this is because there is no commercial software to solve this problem. In addition, there is still the necessary time to plan the route, in order to minimize fuel consumption and ensure the transport of priority items. In these circumstances, it is of great importance to find a method that speeds up the load planning and the definition of the flight itinerary.

@@

Throughout this text, we call {\it Packed Contents}\/ (PC) a set of items of same destination stacked on a pallet and covered with a restraining net (see Figure \ref{fig:larger2}). It is considered unique, having the same attributes of its components, whose values are the sum of individual scores, weights and volumes. To ensure accuracy in pickup and delivery operations, a PC must remain on board until its destination.

@@

We consider real scenarios with an aircraft with payload of $75,000 kg$, with enough fuel for a range of $4,400km$. The aircraft layout is presented in Figure \ref{fig:larger}, where the pallets positions are identified by $p_i$, $1\leq i \leq 18$.

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.4]{Images/consolidated.jpg}
	\caption{A PC on a 463L pallet inside a {\it Boeing C-17}}
	\small\textsuperscript{Source: From Wikimedia Commons, the free media repository}
	\label{fig:larger2}
\end{figure}

@@

Table \ref{tab:larger} shows the parameters of this aircraft. $p_i$\/ are pallets, $1\leq i \leq 18$, with weight $W_i$\/ and $V_i$\/ volume limits, respectively. $D_i^{long}$\/ and $D_i^{lat}$\/ are, respectively, the longitudinal and latitudinal distances of each pallet centroids to the CG of aircraft along both axes. These distances will be used in the calculation of the torque referring to the items allocated on each pallet. In this aircraft, as the ramp have an inclination of $25°$, we made the necessary corrections in $D_i^{long}$, $W_i$\/ and $V_i$\/ of the corresponding pallets.

@@

This aircraft spends $c_d$\/ dollars per kilometer flown, and can carry up to $W_{max}$\/ of cargo distributed on the pallets. The {\it fuel consumption rate}\/ $c_g$\/ is the percentage limit of cost increase  due to the CG deviation on the longitudinal axis of the aircraft. $c_g$\/ depends on the characteristics of the aircraft, and is an arbitrated value for this work that should be better evaluated in future research. It is important to consider that the $c_g$\/ tends to zero as the aircraft attitude tends to level. As the CG deviation varies from $0$\/ to $limit^{CG}_{long}$, the {\it fuel cost}\/ increase varies from $0$\/ to $c_g$.

@@

Let $S_K = \{s: \{1, \dots, K\} \rightarrow \{1, \dots, K\} \}$\/ be the set of $K!$ permutations $\pi$, which correspond to all possible tours (or itineraries) that have node $0$\/ as origin and end, passing through the other $K$\/ nodes. Let $\pi_k$\/ be $k^{th}$\/ node of the tour $\pi$, $1 \leq k \leq K$. In this way, the tour $\pi$\/ is described as $\{0,\pi_1,\ldots,\pi_K, 0\}$, {\color{blue}$K > 1$}. For ease of notation, we can define $\pi_0 = \pi_{K+1} = 0$.

We consider K > 1 because the case K = 1 has a single and trivial solution for the tour.

@@

$\tilde{s}_\pi$\/ \ref{eq:scores} is the sum of the scores of the items loaded on the aircraft throughout the tour $\pi$.

@@

$\tau_{\pi_k}$\/ \ref{eq:tau} is the longitudinal torque applied by the loaded pallets, in the node ${\pi_k}$, in proportion relative to the highest torque supported by the aircraft.

@@

In this way, we can calculate the cost $\tilde{c}_\pi$\/ \ref{eq:costs}, referring to the total fuel consumed in the tour $\pi$\/ due to the distances traveled and the CG longitudinal deviations in each pallet.

@@

The sets of nodes not yet visited are defined in \ref{eq:pdp11} and \ref{eq:departing}.

@@

At the beginning of the tour $\pi$, there are no packed contents \ref{eq:pdp13}.

@@

$\epsilon_{\pi_k}^t$\/ \ref{eq:LatItem} and $\epsilon_{\pi_k}^a$\/ \ref{eq:LatPacked} are the lateral torque applied by the loaded pallets, in the node ${\pi_k}$, in proportion relative to the highest torque supported by the aircraft. For ease of notation, $\epsilon_{\pi_k}^t$\/ corresponds to the items, and $\epsilon_{\pi_k}^a$\/ to the packed contents.

@@

Finally, we can consider the constraints on each node $\pi_k$:

\begin{itemize}
	\item The longitudinal \ref{eq:torqlong} and the lateral \ref{eq:torqlat} torques must be within the limits of the aircraft.
	\item The items allocated to each pallet cannot exceed its weight \ref{eq:app2} and volume \ref{eq:app3} limits.
	\item At most, each item is associated with a single pallet \ref{eq:app4};
	\item Packet contents that have not yet reached their destination must remain on board \ref{eq:app5}.
	\item Items allocated on the same pallet must have the same destinations. In this case, we need two constraints: \ref{eq18} and \ref{eq19}. If $X_{ia}^{\pi_k} = X_{ib}^{\pi_k} = 1$, both constraints require that $to_a = to_b$; otherwise these constraints have no effect.
	\item If there is a packed content on the pallet, it must also have the same destination as the other items. Similarly, we use two constraints: \ref{eq20} and \ref{eq21}.
\end{itemize}

@@

Throughout our research, we have thoughtfully described the ACLP+RPDP model in standard MIP format and found that no MIP solver can handle its practical cases in a feasible time. Thus, as ACLP+RPDP is highly complex, involving four intractable subproblems (APP, WBP, PDP and TSP), our strategy will be to focus only on real cases and to develop quick node-by-node solutions, not necessarily optimal, that allow to build a complete tour.

@@

On the other hand, we know that transport aircraft generally have a few dozen pallets, the flight plan has less than $6$\/ nodes, and each node has hundreds of items to be shipped. Under these circumstances, we can adopt two more strategies:

@@

\item Considering the number of destinations less than the number of pallets ($K < m$), we can preset the destinations of the pallets in each node. In this way, we will reserve a number of pallets proportional to the volume demanded by each destination at the shipping node. We could have used another criterion, but it was observed in the experiments that volume is more constrictive in airlift.

\item If the number of nodes is small and we have quick node-by-node solutions, we have the possibility to check the solution corresponding to the tour with the shortest total distance, and even test all possible tours, selecting the one that provides the best value for the objective function.

@@

\begin{algorithm}[H]
	\caption{ACLP+RPDP solving}  \label{alg:main}
	\begin{algorithmic}[1]

@@

		\State Let aircraft and $M$\/ be according to Table \ref{tab:larger}  \label{main:M} \Comment{Aircraft parameters}

		\State Let $K$, $L$\/ and $C$ be according to $scenario$\/ and Table \ref{tab:costs} \label{main:KLC} \Comment{Nodes, distances and costs}

		\State Let $\pi_{TSP}$\/ be the tour with the shortest distance traveled

@@

			\State {\color{blue} $f_{\pi_{dir}} \gets SolveTour(\pi_{dir}, L, M, C, N, method, time)$ } \label{main:fdirect}

		    \State {\color{blue} $f_{\pi_{rev}} \gets SolveTour(\pi_{rev}, L, M, C, N, method, time)$ } \label{main:freverse}

			\State {\color{blue} $answer1[scenario,surplus,method] \gets max( f_{\pi_{dir}}, f_{\pi_{rev}}  )$ } \label{main:a1}

@@

				\State $f_{\pi} \gets SolveTour(\pi, L, M, C, N, method, time/(K+1)!)$ \Comment{$f_{\pi}$\/ is a solution to the tour $\pi$} \label{main:f2}

@@

		\Return $answer$

@@

{\color{blue} In line \ref{main:fdirect}, $SolveTour$ solves the lowest-cost tour, and in line 8, its reverse. }

The aircraft parameters are in Table \ref{tab:larger} (line \ref{main:M}). In this algorithm, we use five values for $scenario$, according to Table \ref{tab:scenarios}, which defines $K$\/ and the set $L$\/ of nodes (line \ref{main:KLC}).

\vspace{2.0mm}
\begin{table}[H]
	\centering
	\caption{Testing scenarios}  \label{tab:scenarios}
	\begin{tabular}{c c c c}
		\toprule
		{\bf Scenario} & {$K$} & {$L$} & {Shortest tours} \\		
		\midrule
		1 & 2    & \{$1,2$\}           & 0 1 2 0 \\
		2 & 3    & \{$1,2,3$\}         & 0 1 2 3 0 \\
		3 & 4    & \{$1,2,3,4$\}       & 0 4 1 2 3 0 \\
		4 & 5    & \{$1,2,3,4,5$\}     & 0 4 1 2 5 3 0 \\
		5 & 6    & \{$1,2,3,4,5,6$\}   & 0 4 1 2 6 5 3 0 \\
		\bottomrule
	\end{tabular}
\end{table}

@@

$time$\/ is a run time limit, which will be distributed among the node-by-node solutions (lines \ref{main:f1} and \ref{main:f2}). Each tour consists of $K+1$\/ trips, and the total number of tours is $K!$.

$method$\/ corresponds to a MIP solver or a heuristic that we will present in subsection \ref{methods}, where $SolveTour$\/ is the node-by-node solution.

The results corresponding to the shortest tour are stored in $answer1$\/ (line \ref{main:a1}), and those obtained by testing all tours are stored in $answer2$\/ (line \ref{main:a2}).

Next, we will present two subsections: in the first we explain how $SolveTour$\/ is executed, presetting the destinations of the pallets. In the second we will present the heuristic developed for node-by-node solutions.

@@

As we commented in the previous subsection, we will adopt the strategy of presetting the destinations of each pallet throughout the tour. This is feasible in practical cases, where $K<m$. For this, each pallet $i$\/ also has a field $T^k_i$, $0\leq k \leq K$, which stores its next destination after being loaded in node $k$. For this reason, $T^k_i \in L_k$, $1 \leq i \leq m$, $0\leq k \leq K$.

$SolveTour$\/ is described in Algorithm \ref{alg:tour}, where $\pi$\/ is a permutation of the nodes (excluding the base) that defines the order of visits in this tour, $method$\/ corresponds to a MIP solver or a heuristic for solving the node-by-node problems, and $time$\/ is the run time limit that will be distributed among the $K+1$\/ legs of the tour.

\begin{algorithm}[H]
	\caption{Procedure to solve tour $\pi$ with $method$ (MIP solver or heuristic)}  \label{alg:tour}
	\begin{algorithmic}[1]

@@

		\For {$k \gets 0$ to $K$} \label{tour:loop1}	
		\State $L_{\pi_k} \gets L - \{\pi_0,\pi_1,\ldots,\pi_k\}$  \label{tour:lk1}	\Comment{$\pi_k$ is the current node}	
		\For{$i \gets 1$ to $m$}
		\State $T_i^{\pi_k} \gets -1$ \label{tour:-1} \Comment{Unset the pallet destinations}
		\EndFor

@@

		\State $G_2 \gets SolveNode(method, \pi_k, G_1, time/(K+1))$ \label{tour:node}

@@

When the aircraft is at the base, the initial graph $G_1$\/ is empty because it has no packed content \ref{tour:g11}. Otherwise, $UpdatePacked$\/ (line \ref{tour:dest}) returns the set of packed contents that have not yet reached their destination and remain on board, rearranging them on the pallets to minimize CG deviation. This allocation is stored in graph $G_1$\/ (line \ref{tour:g12}).

@@

\begin{algorithm}[H]
	\caption{Procedure to update the PC that remain boarded on node $\pi_k$}  \label{alg:cons}
	\begin{algorithmic}[1]

@@

		\State $E_{Q_{\pi_k}} \gets MinCGDeviation(E_{Q_{\pi_k}})$ \label{cons:minCG}

@@

$MinCGDeviation$\/ (line \ref{cons:minCG}) is run through a MIP solver with the equations \ref{eq:torque}, \ref{eq:embarked} and \ref {eq:one}, to relocate the packed contents on the pallets, minimizing torque and ensuring that they all remain on board, one packed content on each pallet. As there are few variables, the MIP solver returns an allocation $E_{Q_{\pi_k}}$\/ in less than 30 milliseconds. Finally, the destination of each pallet with a packed content is updated (lines \ref{cons:Ybegin}-\ref{cons:Yend}).

\vspace{3mm}
$MinCGDeviation$ objective function:

\begin{equation} \label{eq:torque}
	\mbox{minimize}\ \Big | \sum_{i=1}^{m} \sum_{q=1}^{m_{\pi_k}} Y^k_{iq} \times w_q \times D_i^{long}  \Big |
\end{equation}

\vspace{3mm}
s.t.:

@@

\begin{algorithm}[H]
	\caption{Procedure to set pallets destination based on the items to be embarked on node $\pi_k$}  \label{alg:dest}
	\begin{algorithmic}[1]

@@

			\For{$x \in L_{\pi_k}$}		
				\State $vol_x \gets  0$  \label{dest:vector1}
			\EndFor

@@

						\If{$np < needed$ {\bf and} $T_i^{\pi_k} = -1$}
							\State $T_i^{\pi_k} \gets x$
							\State $np \gets np + 1$
						\EndIf

@@

$ScoreAndDeviation$\/ is described in Algorithm \ref{alg:eval}, which evaluates the allocation graph generated by $SolveNode$, returning the corresponding score and CG deviation.

\begin{algorithm}[H]
	\caption{Procedure to calculate the score and the relative torque deviation of graph $G$\/ in node $\pi_k$}  \label{alg:eval}
	\begin{algorithmic}[1]

@@

		\State $s \gets 0$
		\For{$i \gets 1$ to $m$}
			\State $\tau_i \gets 0$
		\EndFor
		\For{$i \gets 1$ to $m$}	\label{eval:loop1}

@@

		\State $\tau \gets  \sum_{1}^{m} \tau_i / (W_{max} \times limit^{CG}_{long})$ \label{eval:torque}

@@

In this subsection we present two implementations of $SolveNode(method,\ \pi_k,\ G)$, where  $\pi_k$\/ is the current node and $G$\/ is the allocation graph of the PC that remain on board at node $\pi_k$. These implementations correspond to the two possible values of the $method$\/ parameter: $MIP$\/ ({\it Gurobi}) or $Shims$\/ (a heuristic).

\subsubsection{MIP solver}
\label{solver}

To find a viable solution for ACLP+RPDP, the strategy adopted in $SolveTour$\/ previously defines the values of some variables. Concretely, all permutations between the nodes are tested, the set of nodes to be visited is updated, the PC that remain on board are reallocated to minimize the CG deviation, and the pallets destination is determined according to the volume of items available for shipment. In this way, the mathematical modeling for $SolveNode(MIP,\ \pi_k,\ G)$\/ becomes simpler, which finds an allocation of available items in node $\pi_k$\/ using previously defined values of $L_{\pi_k}$, $to_i^{\pi_k}$, and $a^{\pi_k}_q$.

This simplified modeling is described in the equations \ref{eq:obj2} to \ref{eq:if2}, which maximizes the score of the $n_{\pi_k}$ items to be shipped, maintaining load balancing.

The variables $X_{ij}$ and $Y_{iq}$ define the sets of edges $E_{N_{\pi_k}}$ and $E_{Q_{\pi_k}}$, respectively, which will be added to the graph $G$.

\vspace{3mm}
\textbf{Objective function:}

\begin{equation} \label{eq:obj2}
	maximize\ \ \tilde{s} / \tilde{c}
\end{equation}

\vspace{3mm}
\textbf{Where:}

@@

\begin{equation} 
	W_{max} = \sum_{i=1}^{m} W_i
\end{equation}

\begin{equation} 
	\tau^{\pi_k} = \sum_{i=1}^{m} \Big [ D_i^{long} \times (\sum_{j=1}^{n_{\pi_k}} X_{ij} \times w_j +  \sum_{q=1}^{m_{\pi_k}} Y_{iq} \times w_q) \Big ] \Big / W_{max} \times limit^{CG}_{long}
\end{equation}

\vspace{3mm}
In Equation \ref{eq:costs2}, the first parcel denotes the fuel cost from the base until the last node, and the second the fuel cost in the last leg, from the last node back to the base. Note that $c_g$ is the fuel consumption potential increase due to the loaded aircraft CG deviation in the node ($\tau_{\pi_k}$).

\begin{equation} \label{eq:costs2}
	\tilde{c} = \sum_{k=0}^{K} \Big [ c_{\pi_k, \pi_{k+1}}\times(1+c_g\times|\tau_{\pi_k}|) \Big ]
\end{equation}

\vspace{3mm}
\textbf{s.t.:}

{\color{blue} It was observed in the experiments that the lateral torque was always in the order of magnitude -3. So we did not consider these constraints.}

\begin{equation}
	|\tau_{\pi_k}| \leq 1
\end{equation}

\begin{equation} 
	\sum_{i=1}^{m} (\sum_{j=1}^{n_{\pi_k}} X_{ij} \times w_j + \sum_{q=1}^{m_{\pi_k}} Y_{iq} \times w_q ) \leq W_{max}
\end{equation}

@@

\subsubsection{Shims heuristic}

Our goal is to find a heuristic that offers a good-quality solution for the node-by-node problem. As we are testing all $K!$\/ tours, its run time is a crucial requirement. Taking this into account, we perform a series of implementations based on known meta-heuristics: {\it Ant Colony Optimization} (ACO) \cite{Dorigo1992, DorigoManiezzoColorni1996}, {\it Noising Method Optimization} (NMO) \cite{CharonHudry1993,CharonHudry2001,Zhan2020}, {\it Tabu Search} (TS) \cite{Glover1986} and {\it Greedy Randomized Adaptive Search Procedure} (GRASP) \cite{FeoResende1989}.

{\color{blue}
These meta-heuristics were implemented in their standard forms. We did not use any specialized features to allow for the exploitation of the problem structure.
}

>>>>

\usepackage{textcomp}

@@

As can be seen, so far Lurkin and Schyns \cite{LurkinSchyns2015} is the only work that simultaneously addresses an air cargo (WBP) and a flight itinerary (PDP) sub-problem. Although it is innovative, strong simplifications were imposed by these authors: in relation to loading, APP was ignored; with regard to routing, it is assumed a predefined tour plan restricted to only two legs. It is important to note that these authors consider an aircraft with two doors, and the minimization of loading and unloading costs at the intermediate node was modeled through a container sequencing problem. Referring directly to this work, Brandt and Nickel \cite[p. 409]{BrandtStefan2019} comment: {\it However, not even these sub-problems are acceptably solved for real-world problem sizes or the models omit some practically relevant constraints}.

@@

However, as reported by Fok and Chun, {\it load planning (...) is usually done in roughly 2 hours before departure, when all the details of the cargo are present.} \cite{fok2004optimizing}. We can infer that this is because adequate commercial software does not exist. In addition, there is still the necessary time to plan the route, in order to minimize fuel consumption and ensure the transport of priority items. In these circumstances, it is of great importance to find a method that speeds up the load planning and the definition of the flight itinerary.

@@

Throughout this text, we call {\it packed content}\/ a set of items of same destination stacked on a pallet and covered with a restraining net (see Figure \ref{fig:larger2}). It is considered a single item, having the same attributes of its components, whose values are the sum of individual scores, weights and volumes. To ensure accuracy in pickup and delivery operations, a packed content must remain on board until its destination.

@@

We consider real scenarios with an single aircraft with payload of $75,000 kg$, with enough fuel for a range of $4,400km$. The aircraft layout is presented in Figure \ref{fig:larger}, where the pallets positions are identified by $p_i$, $1\leq i \leq 18$.

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.4]{Images/consolidated.jpg}
	\caption{A packed content on a 463L pallet inside a {\it Boeing C-17}}
	\small\textsuperscript{Source: From Wikimedia Commons, the free media repository}
	\label{fig:larger2}
\end{figure}

@@

Table \ref{tab:larger} shows the parameters of this aircraft. $p_i$\/ are pallets, $1\leq i \leq 18$, with weight $W_i$\/ and $V_i$\/ volume limits, respectively. $D_i^{long}$\/ and $D_i^{lat}$\/ are, respectively, the longitudinal and lateral distances of each pallet centroids to the CG of aircraft along both axes. These distances will be used in the calculation of the torque referring to the items allocated on each pallet. In this aircraft, as the ramp have an inclination of $25^{\circ}$, we made the necessary corrections in $D_i^{long}$, $W_i$\/ and $V_i$\/ of the corresponding pallets.

@@

This aircraft spends $c_d$\/ dollars per kilometer flown, and can carry up to $W_{max}$\/ of cargo distributed on the pallets. The {\it fuel consumption rate}\/ $c_g$\/ is the percentage limit of cost increase  due to the CG deviation on the longitudinal axis of the aircraft. $c_g$\/ depends on the characteristics of the aircraft, and is an arbitrated value that should be better evaluated in future research. It is important to consider that the $c_g$\/ tends to zero as the aircraft attitude tends to level. As the CG deviation varies from $0$\/ to $limit^{CG}_{long}$, the {\it fuel cost}\/ increase varies from $0$\/ to $c_g$.

@@

Let $S_K = \{s: \{1, \dots, K\} \rightarrow \{1, \dots, K\} \}$\/ be the set of $K!$ permutations $\pi$, which correspond to all possible tours (or itineraries) that have node $0$\/ as origin and end, passing through the other $K$\/ nodes. Let $\pi_k$\/ be $k^{th}$\/ node of the tour $\pi$, $1 \leq k \leq K$. In this way, the tour $\pi$\/ is described as $\{0,\pi_1,\ldots,\pi_K, 0\}$. For ease of notation, we can define $\pi_0 = \pi_{K+1} = 0$.

@@

We have the following calculation equations:

\begin{itemize}
	\item $\tilde{s}_\pi$\/ \ref{eq:scores} is the sum of the scores of the items loaded on the aircraft throughout the tour $\pi$.
	\item $\tau_{\pi_k}$\/ \ref{eq:tau} is the longitudinal torque applied by the loaded pallets at the node ${\pi_k}$, in proportion relative to the highest torque supported by the aircraft.
	\item $\tilde{c}_\pi$\/ \ref{eq:costs} is the cost of to the total fuel consumed in the tour $\pi$\/ due to the distances traveled and the CG longitudinal deviations \footnote{In our experiments, we found that the magnitude of the lateral torque was always very small, and therefore we decided to ignore it in the fuel consumption.} in each pallet.
	\item The sets of nodes not yet visited along the tour $\pi$\/ are defined in \ref{eq:pdp11} and \ref{eq:departing}.
	\item At the beginning of the tour $\pi$, there are no packed contents \ref{eq:pdp13}. 
	\item $\epsilon_{\pi_k}^t$\/ \ref{eq:LatItem} and $\epsilon_{\pi_k}^a$\/ \ref{eq:LatPacked} are the lateral torque applied by the loaded pallets at the node ${\pi_k}$, in proportion relative to the highest torque supported by the aircraft. For ease of notation, $\epsilon_{\pi_k}^t$\/ corresponds to the items, and $\epsilon_{\pi_k}^a$\/ to the packed contents.
\end{itemize}

@@

@@

@@

@@

@@

@@

Finally, we can consider the constraints at each node $\pi_k$:

\begin{itemize}
	\item The longitudinal \ref{eq:torqlong} and the lateral \ref{eq:torqlat} torques must be within the limits of the aircraft.
	\item The items allocated to each pallet cannot exceed its weight \ref{eq:app2} and volume \ref{eq:app3} limits.
	\item At most, each item is associated with a single pallet \ref{eq:app4};
	\item Packed contents that have not yet reached their destination must remain on board \ref{eq:app5}.
	\item Items allocated on the same pallet must have the same destinations. In this case, we need two constraints: \ref{eq18} and \ref{eq19}. If $X_{ia}^{\pi_k} = X_{ib}^{\pi_k} = 1$, both constraints require that $to_a = to_b$; otherwise these constraints have no effect.
	\item If there is a packed content on the pallet, it must also have the same destination as the other items. Similarly, we use two constraints: \ref{eq20} and \ref{eq21}.
\end{itemize}

@@

Throughout our research, we have thoughtfully described the ACLP+RPDP model in standard MIP format and found that no MIP solver can handle its practical cases in a feasible time. Thus, as ACLP+RPDP is highly complex, involving four intractable subproblems (APP, WBP, PDP and TSP), our strategy will be to focus only on {\it real cases}\/ and to develop quick node-by-node solutions, not necessarily optimal, that allow to build a complete tour.

@@

On the other hand, we know that transport aircraft generally have a few dozen pallets, the flight plan has less than $6$\/ nodes, and each node has hundreds of items to be shipped. We also know that missions with fewer nodes are more frequent than longer missions. Under these circumstances, we can adopt two more strategies:

@@

\item We will consider that the number of destinations is greater than 1, to avoid the case of a single and trivial tour, but less than the number of pallets, that is, $1 < K < m$. So we can preset the destinations of the pallets at each node. In this way, we will reserve a number of pallets proportional to the volume demanded by each destination at the shipping node. We could have used another criterion, but it was observed in the experiments that volume is more constrictive in airlift.

\item If we have quick node-by-node solutions, as the number of nodes is small we have the possibility to check the solution corresponding to the tour with the shortest total distance, and even test all possible tours, selecting the one that provides the best value for the objective function.

@@

\begin{algorithm}[H]
	\caption{Solving ACLP+RPDP}  \label{alg:main}
	\begin{algorithmic}[1]

@@

		\State Let aircraft and $M$  \label{main:M} \Comment{Aircraft parameters from Table \ref{tab:larger}}

		\State Let $K$, $L$\/ and $C$ be according to $scenario$ \label{main:KLC} \Comment{Nodes and costs from Tables \ref{tab:costs} and \ref{tab:scenarios}}

		\State Let $\pi_{TSP1}$\/ and $\pi_{TSP2}$  \Comment{The shortest tours from Tables \ref{tab:costs} and \ref{tab:scenarios}}

@@

			\State $f_1 \gets SolveTour(\pi_{TSP1}, L, M, C, N, method, time)$  \label{main:f1}

		    \State $f_2 \gets SolveTour(\pi_{TSP2}, L, M, C, N, method, time)$  \label{main:f2}

			\State $answer1[scenario,surplus,method] \gets max( f_1, f_2  )$  \label{main:a1} \Comment{Best result between the shortest tours}

@@

				\State $f_{\pi} \gets SolveTour(\pi, L, M, C, N, method, time/K!)$ \Comment{$f_{\pi}$\/ is a solution to the tour $\pi$} \label{main:fpi}

@@

		\Return $answer1, answer2$

@@

The aircraft parameters and the set $M$\/ of pallets are in Table \ref{tab:larger}. In this algorithm, we use five values for $scenario$, according to Tables \ref{tab:costs} and \ref{tab:scenarios}, which define the number $K$ of destinations, the set $L$\/ of nodes, the costs $C$\/, and the shortest tours $\pi_{TSP1}$\/ and $\pi_{TSP2}$.

\vspace{2.0mm}
\begin{table}[H]
	\centering
	\caption{Testing scenarios}  \label{tab:scenarios}
	\begin{tabular}{c c c c c}
		\toprule
		{\bf Scenario} & $K$ & $L$ & $\pi_{TSP1}$ & $\pi_{TSP2}$ \\		
		\midrule
		1 & 2    & \{$0,1,2$\}           & 0 1 2 0  & 0 2 1 0 \\
		2 & 3    & \{$0,1,2,3$\}         & 0 1 2 3 0 & 0 3 2 1 0\\
		3 & 4    & \{$0,1,2,3,4$\}       & 0 4 1 2 3 0 & 0 3 2 1 4 0 \\
		4 & 5    & \{$0,1,2,3,4,5$\}     & 0 4 1 2 5 3 0 & 0 3 5 2 1 4 0 \\
		5 & 6    & \{$0,1,2,3,4,5,6$\}   & 0 4 1 2 6 5 3 0 & 0 3 5 6 2 1 4 0 \\
		\bottomrule
	\end{tabular}
\end{table}

@@

$time$\/ is a run time limit, which will be distributed among the tours (lines \ref{main:f1}, \ref{main:f2} and \ref{main:fpi}). $method$\/ corresponds to a MIP solver or a heuristic to the node-by-node solution $SolveTour$, which will be presented in subsection \ref{methods}.

The best results corresponding to the shortest tours are stored in $answer1$\/ (line \ref{main:a1}), and those obtained by testing all $K!$\/ tours are stored in $answer2$\/ (line \ref{main:a2}).

Next, we will present two subsections: in the first we explain how $SolveTour$\/ is executed, presetting the destinations of the pallets. In the second we will present the heuristics developed for node-by-node solutions.

@@

As we commented in the previous subsection, we will adopt the strategy of presetting the destinations of each pallet throughout the tour. This is feasible in practical cases, where $1<K<m$. For this, each pallet $i$\/ also has a field $T^k_i$, $0\leq k \leq K$, which stores its next destination after being loaded at node $k$. For this reason, $T^k_i \in L_k$, $1 \leq i \leq m$, $0\leq k \leq K$.

$SolveTour$\/ is described in Algorithm \ref{alg:tour}, where $\pi$\/ is a permutation of the nodes (excluding the base) that defines the order of visits in this tour, $method$\/ corresponds to a MIP solver or a heuristic for solving the node-by-node problems, and $time$\/ is the runtime limit that will be distributed among the $K+1$\/ legs of the tour.

\begin{algorithm}[H]
	\caption{Solving the tour $\pi$ with $method$}  \label{alg:tour}
	\begin{algorithmic}[1]

@@

		\For {$k \gets 0$ to $K$} \label{tour:loop1}	
		\State $L_{\pi_k} \gets L - \{\pi_0,\pi_1,\ldots,\pi_k\}$  \label{tour:lk1}	\Comment{$\pi_k$ is the current node}	
		\State $T_i^{\pi_k} \gets -1$, $1 \leq i \leq m$ \label{tour:-1} \Comment{Unset the pallet destinations}

@@

		\State $G_2 \gets SolveNode(method, \pi_k, G_1, time/(K+1))$ \Comment{Each tour has $K+1$\/ legs} \label{tour:node}

@@

When the aircraft is at the base, the initial graph $G_1$\/ is empty and there is no packed contents \ref{tour:g11}. Otherwise, $UpdatePacked$\/ (line \ref{tour:dest}) returns the set of packed contents that have not yet reached their destination and remain on board, rearranging them on the pallets to minimize CG deviation. This allocation is stored in graph $G_1$\/ (line \ref{tour:g12}).

@@

\begin{algorithm}[H]
	\caption{Updating the packed contents that remain boarded at the node $\pi_k$}  \label{alg:cons}
	\begin{algorithmic}[1]

@@

		\State $E_{Q_{\pi_k}} \gets MinCGDeviation(E_{Q_{\pi_k}})$ \Comment{Using a MIP solver} \label{cons:minCG}

@@

$MinCGDeviation$\/ (line \ref{cons:minCG}) relocates the packed contents on the pallets, minimizing torque and ensuring that they all remain on board, one packed content on each pallet. It is run through a MIP solver with objective function \ref{eq:torque} and constraints \ref{eq:embarked} and \ref{eq:one}. As there are few variables, $E_{Q_{\pi_k}}$\/ is obtained in less than $30$\/ milliseconds. Finally, the destination of each pallet with a packed content is updated (lines \ref{cons:Ybegin}-\ref{cons:Yend}).

\begin{equation} \label{eq:torque}
	\min f =  \Big | \sum_{i=1}^{m} \sum_{q=1}^{m_{\pi_k}} Y^k_{iq} \times w_q \times D_i^{long}  \Big |
\end{equation}

@@

\begin{algorithm}[H]
	\caption{Setting pallets destination based on the items to be embarked at the node $\pi_k$}  \label{alg:dest}
	\begin{algorithmic}[1]

@@

			\State $vol_x \gets  0$, $x \in L_{\pi_k}$  \label{dest:vector1}

@@

						\If{($np < needed$) {\bf and} ($T_i^{\pi_k} = -1$)}
							\State $T_i^{\pi_k} \gets x$
							\State $np \gets np + 1$
						\EndIf

@@

$ScoreAndDeviation$, described in Algorithm \ref{alg:eval}, evaluates the allocation graph $G$\/ generated by $SolveNode$\/ at node $\pi_k$\/ and returns the corresponding score and CG deviation.

\begin{algorithm}[H]
	\caption{Calculating the score and the relative torque deviation of the graph $G$\/ at the node $\pi_k$}  \label{alg:eval}
	\begin{algorithmic}[1]

@@

		\State $s \gets 0$
		\State $\tau_i \gets 0$, $1 \leq i \leq m$
		\For{$i \gets 1$ to $m$}	\label{eval:loop1}

@@

		\State $\tau \gets  \sum_{i=1}^{m} \tau_i / (W_{max} \times limit^{CG}_{long})$ \label{eval:torque}

@@

In this subsection we present two implementations of $SolveNode$\/ algorithm: with a MIP solver and a heuristic.

\subsubsection{With a MIP solver}
\label{solver}

Our strategy adopted in $SolveTour$\/ previously defines the values of some variables: the set of nodes to be visited is updated, the packed contents that remain on board are reallocated to minimize the CG deviation, and the pallets destinations are determined according to the volume of items available for shipment.

In this way, the mathematical modeling for $SolveNode(MIP,\pi_k,G, time)$\/ becomes simpler, which finds an allocation of available items at the node $\pi_k$\/ using previously defined values of $L_{\pi_k}$, $T_i^{\pi_k}$, and $a^{\pi_k}_q$. Thus, we use a MIP solver with runtime limit $time$\/ at the node $\pi_k$\/ to maximize the objective function \ref{eq:obj2}, with the calculus equations \ref{eq:newf} to \ref{eq:costs2}, subject to the constraints \ref{eq:tau2} to \ref{eq:ifB}. The binary variables $X_{ij}$\/ and $Y_{iq}$\/ define the sets of edges $E_{N_{\pi_k}}$\/ and $E_{Q_{\pi_k}}$, respectively, included in the graph $G$.

\begin{equation} \label{eq:obj2}
	\max f= \tilde{s} / \tilde{c}
\end{equation}

@@

\begin{equation} 
	\tau_{\pi_k} = \sum_{i=1}^{m} \Big [ D_i^{long} \times (\sum_{j=1}^{n_{\pi_k}} X_{ij} \times w_j +  \sum_{q=1}^{m_{\pi_k}} Y_{iq} \times w_q) \Big ] \Big / W_{max} \times limit^{CG}_{long}
\end{equation}

\begin{equation} \label{eq:costs2}
	\tilde{c} =  c_{\pi_k, \pi_{k+1}}\times(1+c_g\times|\tau_{\pi_k}|)
\end{equation}

\begin{equation} \label{eq:tau2}
	|\tau_{\pi_k}| \leq 1
\end{equation}

@@

The constraints \ref{eq:ifA} and \ref{eq:ifB} are equivalent to $X_{ij} =1$\/ if $to_j = T_i^{\pi_k}$, and $X_{ij} =0$\/ otherwise .

\subsubsection{With Shims heuristic}
